A molecular dynamics study of the solvation of \ce{CO_2}, water, and some organic compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-05-04 ionic_liquids.tex
revision

\begin{abstract}
In this paper we calculate solvation free enegies in ionic liquids and predict macroscopic properties required in the design of chemical processes. Carbon dioxide produced in combustion processes is one of the main sources of anthropogenic generation of greenhouse gases. On the other hand, the potential of \ce{[emim][B(CN)_4]} for its capture has recently been reported. For comparison, the solvents \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what was observed experimentally, namely the pressure required to obtain the same molality of \ce{CO_2} is notably lower in \ce{[emim][B(CN)_4]}. The infinite-dilution activity coefficients of hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in literature.
\end{abstract}

@@

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of the anion \ce{[B(CN)_4]^-]}, whether special interactions with \ce{CO_2} are settled. In this sense, Molecular Simulation provides with a powerful tool for studying solvation processes. To our knowledge, only one study in the literature reports the free energy of solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} \cite{Liu_2014_1} which, in addition, is overestimated. This motivated us to carry out a comparison of different force fields of \ce{[emim][B(CN)_4]} in terms of the free energy of solvation and, indeed,  it was possible to select a model that successfully predicts the value of the Henry constant of \ce{CO_2}. For comparison, the solvents \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what was observed experimentally, that is, a considerably lower pressure is required in \ce{[emim][B(CN)_4]} to obtain the same molality of \ce{CO_2}. The infinite-dilution activity coefficients of solutes hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in the literature.

@@

In the following, the subscripts ``$s$'' and ``$\text{IL}$'' refer to the solute and the solvent (ionic liquid), respectively. For a system containing $N_{\text{IL}}$ solvent molecules and $N_{\text{s}}$ solute molecules at fixed pressure (P) and temperature (T), the difference in free energy between the states associated to $\lambda$ = 1 and $\lambda$ = 0 is defined as
\begin{equation}
\Delta G_{\text{sim}} \equiv k_b T \ln \frac{\bigtriangleup (N_s^l,N_{LI},P,T,\lambda = 1)}{\bigtriangleup (N_s^l,N_{LI},P,T,\lambda = 0)}, 
\end{equation}
where $\bigtriangleup$ denotes the isobaric-isothermal partition function. Shirts~\textit{et al}.~\cite{Shirts_2003} presented a detailed mathematical derivation allowing to relate $\Delta G_{\text{sim}}$ to what is experimentally known as the solvation free energy and obtained the following expression: 
\begin{equation}
\begin{split}
\label{eq:free_solv}
 \Delta G_{\text{solv}}& =  k_b T \ln \left( \frac{P_s}{c_s^{\ast} k_b T} \right)\\ &= \Delta G_{\text{sim}} - k_bT \ln \left( \frac{V^{\ast}}{V_1} \right) ,
\end{split}
\end{equation}
where $k_b$ is the Boltzmann's constant, $V^{\ast}$ and $V_1$ refer to the mean volumes of the systems specified by ($N_s^l-1$,$N_{\text{LI}}$,P,T) and ($N_s^l$,$N_{\text{LI}}$,P,T), respectively; $c_s^{\ast}$ = $N_s^l/V_1$ is the numerical concentration of the solute in the liquid phase and $P_s$ is the partial pressure of the solute in the gas phase. It should be noted that in the derivation of Eq.~\ref{eq:free_solv}, the gas phase was treated as an ideal gas.

For small solutes, such as those considered in this work, the magnitude of the term $V^{\ast}/V_1$ is typically on the order of the free energy uncertainties~\cite{Shirts_2003}. Hence, that term could be neglected and will be omitted hereafter.

\subsection*{Henry constant}
If both $P_s$ and $c_s^{\ast}$ were not large, we can employ Eq.~\ref{eq:free_solv} to predict Henry's law constant $\text{K}_s$, which is given by~\cite{Prausnitz}
\begin{equation}
\text{K}_s =\frac{P_s}{x_s}.
\end{equation}
Therefore, by simply expressing Eq.~\ref{eq:free_solv} in terms of the molar fraction $x_s$ and considering MW (molecular weight of the solution) = $\text{MW}_{\scaleto{LI}{4pt}}$ as well as $\rho$ (mass density of the solution) = $\rho_{\scaleto{LI}{4pt}}$, we obtain that
\begin{equation}
\label{eq:henry_eq}
K_s = \tilde{\rho}_{\scaleto{LI}{4pt}} R T \exp(\beta \, \Delta G_{\text{sim}}),
\end{equation}
where $R$ is the gas constant.

@@

Using standard thermodynamic relations~\cite{Tester}, we write the activity coefficient $\gamma_s$ in terms of residual chemical potentials, as follows
\begin{equation}
\label{eq:gamma}
\ln \gamma_s = \frac{1}{RT} (\mu_s^R - \mu_s^{R,0}),
\end{equation}
where $\mu_s^R$ and $\mu_s^{R,0}$ are the residual chemical potentials of the solute in the solution and in the pure solute, respectively. It is worth noting that we have used the definition $\mu_s^{R}$ (P,T) = $\mu_s$ (P,T) - $\mu_s^{ig}$ (P,T), where the superscript $ig$ refers to a solution of ideal gases. Also, recall that $\mu_s^{ig} (P,T)$ = $\mu_s^{ig,0} (P,T)$ + $RT \ln x_s$, so we have that
\begin{equation}
\mu_s^R = \mu_s^l(N_s^l,N_{LI},P,T) - \mu_s^{ig,0}(P,T) - k_b T \ln x_s. 
\end{equation}
Replacing $\mu_s^{l}$ and $\mu_s^{ig,0}$ by the corresponding expressions derived in Ref.~\cite{Shirts_2003} yields
\begin{equation}
\label{eq:mu_res}
\mu_s^R = \Delta G_{\text{sim}} - k_b T \ln \left( \frac{V^{\ast} P}{N_s^l k_b T} \right) -  k_b T \ln x_s.
\end{equation}
Of course, an analogous equation holds for $\mu_s^{R,0}$:
\begin{equation}
\label{eq:mu_res_0}
\mu_s^{R,0} = \Delta G_{\text{sim}}^{0} - k_b T \ln \left( \frac{V^{\ast,0} P}{N_s^{l,0} k_b T} \right).
\end{equation}

@@

For the ILs of interest to the present investigation, high concentrations of \ce{CO_2} are found at pressure values for which, in principle, the ideal gas assumption may be not valid. In this case, we shall correct the ideal gas contribution to the chemical potential, as follows
\begin{equation}
\label{eq:mu_gas_real}
\mu^{g}_s = \mu^{ig}_s + RT \ln \phi_s,
\end{equation}
where $\phi_s$ is the fugacity coefficient of the solute. This equation, together with the fact that $f^{g}_s$ = $P_s \phi_s$, are used to evaluate $\mu^{l}_s$ = $\mu^{g}_s$, which yields
\begin{equation}
\label{eq:fgas_d}
f^{g}_s = \exp  \left( \beta \Delta G_{\text{sim}} \right) +  k_b T c_s^{\ast}, 
\end{equation}
where $f^{g}_s$ is the fugacity of the solute in the gas phase. Having calculated $f^{g}_s$, $P_s$ could be predicted using the following equation
\begin{equation}
\label{eq:fgas}
f^{g}_s = P_s \exp\left[ \frac{B_s(T) P_s}{R T} \right],
\end{equation}
where $B_s$ corresponds to the second virial coefficient. Here we employ the correlation for \ce{CO_2} introduced by Holste~\textit{et al}~\cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0$ = 23.02991 cm$^3$/mol, $B_1$ = $-2.455297$ $\times$ 10$^3$ (Kcm$^3$/mol) and $B_2$ = -1.22675$\times$10$^7$ (K$^2$cm$^3$/mol). Actually, the prediction of $P_s$ should be iterative and in this work we propose the following scheme: (i) run an NPT simulation, (ii) calculate $\Delta G_{\text{sim}}$ and $f^{g}_s$ via Eq.~\ref{eq:fgas_d}, (iii) employ $f^{g}_s$ in Eq.~\ref{eq:fgas} to compute $P_s$ (iv) if $P_s$ $\neq$ P, repeat the above steps using $P_s$ as the specified pressure in a new NPT simulation. This is done until $P_s$ = P.

@@

The versatility of ILs for different applications is in part due to the fact that their physical properties strongly depend on the cation-anion combination. On the other hand, this characteristic also explains why the desirable transferability of force fields is particularly difficult to attain. Indeed, it is a common practice to extract the potential parameters corresponding to intramolecular (bond stretching, angle bending, torsion rotation) and Lennard-Jones interactions from a well established force field and then derive the partial atomic charges. This procedure is, however, inaccurate as in conventional force fields the determination of the atomic charges precedes the dihedrals parametrization. Another ad-hoc strategy, which intends to crudely incorporate polarization effects, is to scale the partial charges uniformly with a factor which normally takes values in the range 0.8-0.9. This practice has proven to be effective in improving the predictions of dynamic properties~\cite{Schr_der_2012}.

In this paper, we consider four models of \ce{[emim][B(CN)_4]}. Koller~\textit{et al.}~\cite{Koller_2012} were the first to introduce a specific non-polarizable force field of \ce{[emim][B(CN)_4]}. They tested three variations of a \textit{united-atoms} model, in which the alkyl groups $\text{CH}_3$ and $\text{CH}_2$ of \ce{[emim]^+} are replaced by pseudo-atoms. Among them, the ``FF-3'' model best reproduces the experimental values of density, self-diffusion coefficients and viscosity. In short, both the intramolecular and Lennard-Jones parameters of the cation were taken from the work of Liu~\textit{et al.}~\cite{Liu_2006}, which in turn is based on the AMBER~\cite{Cornell_1995} framework. The intramolecular parameters involving the boron atom were extracted from the DREIDING~\cite{Mayo_1990} force field and the anionic vdW parameters were refined based on the values reported by Price~\textit{et al.}~\cite{Price_2001}. The atomic partial charges were derived using the EA-RESP method~\cite{Basma_2001} from \textit{Post-Hartree-Fock ab-initio} calculations at MP2/6–31G$^\ast$+ level of theory, which automatically resulted in total charges on the ions of $\pm$0.8426e.

The potential parameters of [B(CN)\textsubscript{4}]$^{-}$ derived by Koller~\textit{et al.}~\cite{Koller_2012} were employed by Batista~\textit{et al.}~\cite{Batista_2015} to study aqueous solutions of \ce{[emim][B(CN)_4]}. In this case,  the authors considered the \textit{all-atom} model of \ce{[emim]^+} developed by Cadena and Maginn \cite{Cadena_2006}. The atomic partial charges were obtained using the CHELP procedure \cite{Breneman_1990}, resulting in net charges of $\pm$0.889$e$ on the ions. It should be noted that those force fields consider different scaling factors for 1-4 non-bonded interactions: Cadena and Maginn \cite{Cadena_2006} use values of 1.0 y 0.4 for the Lennard-Jones and Coulombic interactions, respectively, while in Koller \textit{et al.} \cite{Koller_2012} the corresponding factors are 0.5 y 5/6, in accordance with AMBER force field. Batista \textit{et al.} \cite{Batista_2015} employ the same factors as Cadena and Maginn \cite{Cadena_2006} (J.P. Coutinho, personal communication).

Another set of parameters for \mbox{\ce{[emim][B(CN)_4]}} was derived by Liu \textit{et al.} \cite{Liu_2014} in a study that investigates the dynamics of \ce{CO_2} and N\textsubscript{2} in that IL. Parameters corresponding to the intramolecular and Lennard-Jones interactions were extracted from GAFF~\cite{Wang_2004}, while the atomic partial charges were computed using the RESP method~\cite{Bayly_1993} from \textit{DFT}/B3LYP/6-311+G(d) calculations. The net charges of the ions were manually adjusted to a value of $\pm$0.8$e$.

@@

Finally, for [emim][Tf\textsubscript{2}N] we use the model of K\"{o}ddermann~\textit{et al.} \cite{K_ddermann_2007}, which result from refining the force field due to P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}. This refining was characterized by their authors as unusual, given that the LJ parameters were modified, while keeping the partial charges at their original values. This model was recently employed by Kerl \textit{et al.} \cite{Kerl__2017}, who were able to accurately predict Henry's constant for \ce{CO_2} at several temperatures.

@@

The solutes considered here are hexane, benzene, \ce{CO_2}, ethanol, cyclohexane and water. Hexane and benzene were modeled using the OPLS-AA force field~\cite{Jorgensen_1996}, while the TRAPPE potential~\cite{Chen_2001,Potoff_2001} was employed for ethanol and \ce{CO_2}. Cyclohexane was simulated using a rigid-body model~\cite{munoz2015lennard} whose sites only interact via the Lennard-Jones potential and for water we employ the TIP4P/2005 model~\cite{Abascal_2005}, \cite{Vega_2011}.

@@

With the aim of setting up a (more) computationally efficient protocol for the ILs, in this work we assess the possibility of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method~\cite{Fennell2006} for computing the electrostatic interactions. As an alternative to the Ewald schemes~\cite{Ewald_1921,Darden_1993,Hockney_1988}, McCann~\textit{et al.}~\cite{McCann_2013} tested the DSF method for various ILs modeled with the OPLS-AA force fields. They concluded that using $\alpha$ = 0.2 {\AA}$^{-1}$ and a cutoff distance ($r_c$) of 15 {\AA} provides the best energetic agreement with Ewald. In this work, the comparisons are made in terms of the simulated densities and radial distribution functions for which, as will be shown in the next section, $\alpha$ = 0.2 {\AA}$^{-1}$ and $r_c$ = 12 {\AA} resulted in good agreement for the ILs studied here. Those values were also employed for the charged solutes, with the exception of hexane, for which we considered undamped Coulombic interactions.

rigid treatment of the imidazolium ring~\cite{Kobrak_2004,Lynden_Bell_2006}
\cite{Rey_Castro_2007}

When it was possible to treat the ions as rigid bodies, the NPT algorithm of Kamberaj~\textit{et al.}~\cite{Kamberaj_2005} was employed to carry out the MD simulations. Otherwise, the scheme due to Shinoda~\textit{et al}.~\cite{Shinoda2004} was used along with the application of SHAKE~\cite{Ryckaert1977} to constrain the C-H bonds lengths. For both integrators, we employed $\Delta t$ = 1 fs.

The gradual insertion/deletion of charged solutes involves scaling both the van der Waals (vdW) and the electrostatic intermolecular interactions. In order to avoid numerical instabilities associated to the vdW scaling, a standard approach is to employ a soft-core potential with a nonlinear coupling, and in this paper we use the modified Lennard-Jones potential due to Beutler~\textit{et al.}~\cite{Beutler_1994}. The electrostatic interactions, on the other hand, can be handled simply by a linear scaling~\cite{Naden_2015} of the potential, which in this work corresponds to the DSF method~\cite{Fennell2006}.

@@

In this work, we performed independent simulation at each state of $\lambda_{\text{Coul}}$, while the expanded ensemble (EE) method~\cite{Lyubartsev_1992} was applied for sampling the states associated to the $\lambda_{\,\text{vdW}}$ set. In short, an EE is constructed as a weighted sum of subensembles, each assuming a different value of the variable chosen to explore the free energy landscape, which in our case is $\lambda_{\,\text{vdW}}$. This way, the sampling of the EE involves both sampling each subensamble with fixed variables (N,P,T,$\lambda_{\,\text{vdW}}$) and transitions in $\lambda_{\,\text{vdW}}$. Thus, the remarkable feature of the EE method is the possibility of exploring the whole alchemical route in a single simulation. In this work, we perform a sequential sampling of the EE \cite{Chodera_2011_2}, which consists in sampling each subensamble via NPT MD for a determined number of steps $N^{\text{EE}}$ followed with an attemp to change $\lambda_{\,\text{vdW}}$ using MC, and so on. Independent sampling of $\lambda$~\cite{Chodera_2011_2}. In addition, the set of (arbitrary) weights $\eta$ of the subensambles enters in the acceptance criteria of MC, so it is possible to tune $\eta$ in order to enhance the sampling. If differences $\eta_i$ - $\eta_j$ were closed to $\Delta G_{ji}$, we expect an uniform sampling in $\lambda_{\,\text{vdW}}$~\cite{Lyubartsev_1992}. Of course, this can be achieved only iteratively. The EE procedure, together with the \textit{soft-core} potential and the DSF method with linear coupling were all implemented in the LAMMPS software package~\cite{Plimpton1995}, which we employed for carrying out all the MD simulations.

As demonstrated in Ref.~\citenum{Chodera_2011_2}, smaller values of $N^{EE}$ generate more decorrelated samples in each subensemble. In this work, we employ $N^{EE}$ = 5 which, together with the specified $\lambda_{\, \text{vdW}}$ set, resulted to be satisfactory as up to 3 preliminary runs of the EE were enough to determine the weights $\eta$ that yielded, in a final simulation, a ``sufficient'' sampling of each subensamble. By ``sufficient'', we mean that the sampling provides an acceptable error in $\Delta G$. The simulation times for the preliminary and final EE runs were 10 and 40 ns, respectively, and each $\lambda_{\text{Coul}}$ state was simulated for 10 ns. Finally, the MBAR method~\cite{Shirts_2008}, as implemented by Klimovich~\textit{et al.}~\cite{Klimovich_2015}, was applied to estimate $\Delta G_{\text{vdW}}$ and $\Delta G_{\text{Coul}}$ .

>>>>

\begin{abstract}
In this paper we calculate solvation free enegies in ionic liquids and predict macroscopic properties required in the design of chemical processes. Carbon dioxide produced in combustion processes is one of the main sources of anthropogenic generation of greenhouse gases. On the other hand, the potential of \ce{[emim][B(CN)_4]} for its capture has recently been reported. For comparison purposes, the solvents \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what was observed experimentally, namely the pressure required to obtain the same molality of \ce{CO_2} is notably lower in \ce{[emim][B(CN)_4]}. The infinite-dilution activity coefficients of hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in literature.
\end{abstract}

@@

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of the anion \ce{[B(CN)_4]^-]}, whether special interactions with \ce{CO_2} are settled. In this sense, Molecular Simulation provides with a powerful tool for studying solvation processes. To the best of our knowledge, only one study in the literature reports the free energy of solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} \cite{Liu_2014_1} which, in addition, is overestimated. This motivated us to carry out a comparison of different force fields of \ce{[emim][B(CN)_4]} in terms of the free energy of solvation and, indeed,  it was possible to select a model that successfully predicts the value of the Henry constant of \ce{CO_2}. For comparison, the solvents \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what was observed experimentally, that is, a considerably lower pressure is required in \ce{[emim][B(CN)_4]} to obtain the same molality of \ce{CO_2}. The infinite-dilution activity coefficients of solutes hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in the literature.

@@

In the following, the subscripts ``$s$'' and ``$\text{IL}$'' refer to the solute and the solvent (ionic liquid), respectively. For a system containing $N_{\text{IL}}$ solvent molecules and $N_{\text{s}}$ solute molecules at fixed pressure (P) and temperature (T), the difference in free energy between the states associated to $\lambda$ = 1 and $\lambda$ = 0 is defined as
\begin{equation}
\Delta G_{\text{sim}} \equiv k_b T \ln \frac{\bigtriangleup (N_s^l,N_{LI},P,T,\lambda = 1)}{\bigtriangleup (N_s^l,N_{LI},P,T,\lambda = 0)}, 
\end{equation}
where $\bigtriangleup$ denotes the isobaric-isothermal partition function. Shirts \textit{et al}. \cite{Shirts_2003} presented a detailed mathematical derivation allowing to relate $\Delta G_{\text{sim}}$ to what is experimentally known as the solvation free energy and obtained the following expression: 
\begin{equation}
\begin{split}
\label{eq:free_solv}
 \Delta G_{\text{solv}}& =  k_b T \ln \left( \frac{P_s}{c_s^{\ast} k_b T} \right)\\ &= \Delta G_{\text{sim}} - k_bT \ln \left( \frac{V^{\ast}}{V_1} \right) ,
\end{split}
\end{equation}
where $k_b$ is the Boltzmann's constant, $V^{\ast}$ and $V_1$ refer to the mean volumes of the systems specified by ($N_s^l-1$,$N_{\text{LI}}$,P,T) and ($N_s^l$,$N_{\text{LI}}$,P,T), respectively; $c_s^{\ast}$ = $N_s^l/V_1$ is the numerical concentration of the solute in the liquid phase and $P_s$ is the partial pressure of the solute in the gas phase. It should be noted that in the derivation of Eq.~\ref{eq:free_solv}, the gas phase was treated as an ideal gas.

For small solutes, such as those considered in this work, the magnitude of the term $V^{\ast}/V_1$ is typically on the order of the free energy uncertainties \cite{Shirts_2003}. Hence, that term could be neglected and will be omitted hereafter.

\subsection*{Henry constant}
If both $P_s$ and $c_s^{\ast}$ were not large, we can employ Eq.~\ref{eq:free_solv} to predict Henry's law constant $\text{K}_s$, which is given by \cite{Prausnitz}
\begin{equation}
\text{K}_s =\frac{P_s}{x_s}.
\end{equation}
Therefore, by simply expressing Eq.~\ref{eq:free_solv} in terms of the molar fraction $x_s$ and considering MW (molecular weight of the solution) = $\text{MW}_{\scaleto{LI}{4pt}}$ as well as $\rho$ (mass density of the solution) = $\rho_{\scaleto{LI}{4pt}}$, we obtain that
\begin{equation}
\label{eq:henry_eq}
K_s = \tilde{\rho}_{\scaleto{LI}{4pt}} R T \exp(\beta \, \Delta G_{\text{sim}}),
\end{equation}
where $R$ is the gas constant.

@@

Using standard thermodynamic relations \cite{Tester}, we write the activity coefficient $\gamma_s$ in terms of residual chemical potentials, as follows
\begin{equation}
\label{eq:gamma}
\ln \gamma_s = \frac{1}{RT} (\mu_s^R - \mu_s^{R,0}),
\end{equation}
where $\mu_s^R$ and $\mu_s^{R,0}$ are the residual chemical potentials of the solute in the solution and in the pure solute, respectively. It is worth noting that we have used the definition $\mu_s^{R}$ (P,T) = $\mu_s$ (P,T) - $\mu_s^{ig}$ (P,T), where the superscript $ig$ refers to a solution of ideal gases. Also, recall that $\mu_s^{ig} (P,T)$ = $\mu_s^{ig,0} (P,T)$ + $RT \ln x_s$, so we have that
\begin{equation}
\mu_s^R = \mu_s^l(N_s^l,N_{LI},P,T) - \mu_s^{ig,0}(P,T) - k_b T \ln x_s, 
\end{equation}
where $\mu_s^l$ is the chemical potential of the solute in the liquid solution and $x_s$ is the molar fraction. Replacing $\mu_s^{l}$ and $\mu_s^{ig,0}$ by the corresponding expressions derived in Ref. \cite{Shirts_2003} yields
\begin{equation}
\label{eq:mu_res}
\mu_s^R = \Delta G_{\text{sim}} - k_b T \ln \left( \frac{V^{\ast} P}{N_s^l k_b T} \right) -  k_b T \ln x_s.
\end{equation}
Of course, an analogous equation holds for $\mu_s^{R,0}$:
\begin{equation}
\label{eq:mu_res_0}
\mu_s^{R,0} = \Delta G_{\text{sim}}^{0} - k_b T \ln \left( \frac{V^{\ast,0} P}{N_s^{l,0} k_b T} \right).
\end{equation}

@@

For the ILs of interest to the present investigation, high concentrations of \ce{CO_2} are found at pressure values for which, in principle, the ideal gas assumption may be not valid. In this case, we shall correct the ideal gas contribution to the chemical potential, as follows
\begin{equation}
\label{eq:mu_gas_real}
\mu^{g}_s = \mu^{ig}_s + RT \ln \phi_s,
\end{equation}
where $\phi_s$ is the fugacity coefficient of the solute. This equation, together with the fact that $f^{g}_s$ = $P_s \phi_s$, are used to evaluate $\mu^{l}_s$ = $\mu^{g}_s$, which yields
\begin{equation}
\label{eq:fgas_d}
f^{g}_s = \exp  \left( \beta \Delta G_{\text{sim}} \right) +  k_b T c_s^{\ast}, 
\end{equation}
where $f^{g}_s$ is the fugacity of the solute in the gas phase. Having calculated $f^{g}_s$, $P_s$ could be predicted using the following equation
\begin{equation}
\label{eq:fgas}
f^{g}_s = P_s \exp\left[ \frac{B_s(T) P_s}{R T} \right],
\end{equation}
where $B_s$ corresponds to the second virial coefficient. Here we employ the correlation for \ce{CO_2} introduced by Holste \textit{et al} \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0$ = 23.02991 cm$^3$/mol, $B_1$ = $-2.455297$ $\times$ 10$^3$ (Kcm$^3$/mol) and $B_2$ = -1.22675$\times$10$^7$ (K$^2$cm$^3$/mol). Actually, the prediction of $P_s$ should be iterative and in this work we propose the following scheme: (i) run an NPT simulation, (ii) calculate $\Delta G_{\text{sim}}$ and $f^{g}_s$ via Eq.~\ref{eq:fgas_d}, (iii) employ $f^{g}_s$ in Eq.~\ref{eq:fgas} to compute $P_s$ (iv) if $P_s$ $\neq$ P, repeat the above steps using $P_s$ as the specified pressure in a new NPT simulation. This is done until $P_s$ = P.

@@

The versatility of ILs for different applications is in part due to the fact that their physical properties strongly vary with the particular cation-anion combination. On the other hand, this characteristic also explains why the desirable transferability of force fields is particularly difficult to attain. Indeed, it is a common practice to extract the potential parameters corresponding to intramolecular (bond stretching, angle bending, torsion rotation) and Lennard-Jones (LJ) interactions from a well established force field and then derive the partial atomic charges. This procedure is, however, inaccurate as in conventional force fields the determination of the atomic charges precedes the dihedrals parametrization. Another ad-hoc strategy, which intends to crudely incorporate polarization effects, is to scale the partial charges uniformly with a factor which normally takes values in the range 0.8-0.9. This practice has proven to be effective in improving the predictions of dynamic properties \cite{Schr_der_2012}.

In this paper, we consider four models of \ce{[emim][B(CN)_4]}. Koller \textit{et al.} \cite{Koller_2012} were the first to introduce a specific non-polarizable force field of \ce{[emim][B(CN)_4]}. They tested three variations of a \textit{united-atoms} model, in which the alkyl groups \ce{CH_3} and \ce{CH_2} of \ce{[emim]^+} are replaced by pseudo-atoms. Among them, the ``FF-3'' model best reproduces the experimental values of density, self-diffusion coefficients and viscosity. In short, both the intramolecular and LJ parameters of the cation were taken from the work of Liu \textit{et al.} \cite{Liu_2006}, which in turn is based on the AMBER \cite{Cornell_1995} framework. The intramolecular parameters involving the boron atom were extracted from the DREIDING \cite{Mayo_1990} force field and the anionic vdW parameters were refined based on the values reported by Price \textit{et al.} \cite{Price_2001}. The atomic partial charges were derived using the EA-RESP method \cite{Basma_2001} from \textit{Post-Hartree-Fock ab-initio} calculations at MP2/6-31G$^\ast$+ level of theory, which automatically resulted in total charges on the ions of $\pm$0.8426$e$.

The potential parameters of \ce{[B(CN)_4]^-} derived by Koller \textit{et al.} \cite{Koller_2012} were then employed by Batista \textit{et al.} \cite{Batista_2015} to study aqueous solutions of \ce{[emim][B(CN)_4]}. In this case, the authors considered the \textit{all-atom} model of \ce{[emim]^+} developed by Cadena and Maginn \cite{Cadena_2006}. The atomic partial charges were obtained using the CHELP procedure \cite{Breneman_1990}, resulting in net charges of $\pm$0.889$e$ on the ions. It should be noted that those force fields consider different scaling factors for the 1-4 non-bonded interactions: Cadena and Maginn \cite{Cadena_2006} use values of 1.0 y 0.4 for the LJ and Coulombic interactions, respectively; while in Koller \textit{et al.} \cite{Koller_2012} the corresponding factors are 0.5 y 5/6, in accordance with the AMBER force field. Batista \textit{et al.} \cite{Batista_2015} employ the same factors as Cadena and Maginn \cite{Cadena_2006} (J.P. Coutinho, personal communication).

Another set of parameters for \ce{[emim][B(CN)_4]} was derived by Liu \textit{et al.} \cite{Liu_2014} in a study that investigates the dynamics of \ce{CO_2} and \ce{N_2} in that IL. Parameters corresponding to the intramolecular and LJ interactions were extracted from GAFF \cite{Wang_2004}, while the atomic partial charges were computed using the RESP method \cite{Bayly_1993} from \textit{DFT}/B3LYP/6-311+G(d) calculations. The net charges of the ions were manually adjusted to a value of $\pm$0.8$e$.

@@

Finally, for [emim][Tf\textsubscript{2}N] we use the model of K\"{o}ddermann~\textit{et al.} \cite{K_ddermann_2007}, which was derived by refining the force field due to P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}. This refining was characterized by their authors as ``unusual'', given that the LJ parameters were modified, while keeping the partial charges at their original values. This model was recently employed by Kerl \textit{et al.} \cite{Kerl__2017}, who were able to accurately predict Henry's constant of \ce{CO_2} at several temperatures.

@@

The solutes considered here are hexane, benzene, \ce{CO_2}, ethanol, cyclohexane and water. Hexane and benzene were modeled using the OPLS-AA force field \cite{Jorgensen_1996}, while the TRAPPE potential \cite{Chen_2001,Potoff_2001} was employed for ethanol and \ce{CO_2}. Cyclohexane was simulated using a rigid-body model \cite{munoz2015lennard} whose sites only interact via the LJ potential and for water we employ the TIP4P/2005 model \cite{Abascal_2005,Vega_2011}.

@@

With the aim of setting up a (more) computationally efficient protocol for the ILs, in this work we assess the possibility of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method \cite{Fennell2006} for computing the electrostatic interactions. As an alternative to the Ewald schemes \cite{Ewald_1921,Darden_1993,Hockney_1988}, McCann \textit{et al.} \cite{McCann_2013} tested the DSF method for various ILs modeled with the OPLS-AA force fields. They concluded that using $\alpha$ = 0.2 {\AA}$^{-1}$ together with a cutoff distance ($r_c$) of 15 {\AA} provide the best energetic agreement with the Ewald method. In this work, the comparisons are made in terms of the simulated densities and radial distribution functions for which, as will be shown in the next section, $\alpha$ = 0.2 {\AA}$^{-1}$ and $r_c$ = 12 {\AA} resulted in good agreement for the ILs studied here. These values were also employed for the charged solutes, with the exception of hexane, for which we considered undamped Coulombic interactions. The reason for this exception is that we were able to improve the agreement with the experimental solvation free energy of that pure substance.

When it was possible to treat the ions as rigid bodies, the NPT algorithm of Kamberaj \textit{et al.} \cite{Kamberaj_2005} was employed to carry out the MD simulations. Otherwise, the scheme due to Shinoda \textit{et al}. \cite{Shinoda2004} was used along with the application of SHAKE \cite{Ryckaert1977} to constrain the C-H bonds lengths. For both integrators, we employed $\Delta t$ = 1 fs.

The gradual insertion/deletion of charged solutes involves scaling both the van der Waals (vdW) and the electrostatic intermolecular interactions. In order to avoid numerical instabilities associated to the vdW scaling, a standard approach is to employ a soft-core potential with a nonlinear coupling. In this paper we use the modified LJ potential due to Beutler \textit{et al.} \cite{Beutler_1994}. The electrostatic interactions, on the other hand, can be handled simply by a linear scaling of the potential \cite{Naden_2015}.

@@

In this work, we carried out independent simulation at each state of $\lambda_{\text{Coul}}$, while the expanded ensemble (EE) method~\cite{Lyubartsev_1992} was applied for sampling the states associated to the $\lambda_{\,\text{vdW}}$ set. In short, an EE is constructed as a weighted sum of subensembles, each assuming a different value of the variable chosen to explore the free energy landscape, which in our case is $\lambda_{\,\text{vdW}}$. This way, the sampling of the EE involves both sampling each subensamble with fixed variables (N,P,T,$\lambda_{\,\text{vdW}}$) and transitions in $\lambda_{\,\text{vdW}}$. Thus, a remarkable feature of the EE method is the possibility of exploring the whole alchemical route in a single simulation. In this work, we perform a sequential sampling of the EE \cite{Chodera_2011_2}, which consists in sampling each subensamble via NPT MD for a determined number of steps $N^{\text{EE}}$ followed by an attemp to change $\lambda_{\,\text{vdW}}$ using MC, and so on. That change in $\lambda_{\,\text{vdW}}$ is randomly proposed according to an independent sampling scheme \cite{Chodera_2011_2}. In addition, the set of (arbitrary) weights $\eta$ of the subensambles enters in the acceptance criteria of MC, so it is possible to tune $\eta$ in order to enhance the sampling. If differences $\eta_i$ - $\eta_j$ were closed to $\Delta G_{ji}$, we expect an uniform sampling in $\lambda_{\,\text{vdW}}$ \cite{Lyubartsev_1992}. Of course, this can be achieved only iteratively. The EE procedure, together with the \textit{soft-core} potential and the DSF method with linear coupling were all implemented in the LAMMPS software package \cite{Plimpton1995}, which we employed for carrying out all the MD simulations.

As demonstrated in Ref. \citenum{Chodera_2011_2}, smaller values of $N^{EE}$ generate more decorrelated samples in each subensemble. In this work, we employ $N^{EE}$ = 5 which, together with the specified $\lambda_{\, \text{vdW}}$ set, resulted to be satisfactory as up to 3 preliminary runs of the EE were enough to determine the weights $\eta$ that yielded, in a final simulation, a ``sufficient'' sampling of each subensamble. By ``sufficient'', we mean that the sampling provides an acceptable error in $\Delta G$. The simulation times for the preliminary and final EE runs were 10 and 40 ns, respectively; and each $\lambda_{\text{Coul}}$ state was simulated for 10 ns. Finally, the MBAR method \cite{Shirts_2008}, as implemented by Klimovich \textit{et al.} \cite{Klimovich_2015}, was applied to estimate $\Delta G_{\text{vdW}}$ and $\Delta G_{\text{Coul}}$ .
